---
title: "Sampling density of PHIA surveys per polygon"
format: html
editor: visual
---

## Zimbabwe

Import the data

```{r}
#| echo: false
pacman::p_load(rio, tidyverse,sp, sf, raster, ggmap, 
               here, janitor, viridis, terra, tmap)

zim_centroids  <- import(here("data", "raw", "PHIA data", 
                              "Zimbabwe", "zimphia2015centroids.csv")) #centroidid
attr(zim_centroids, 'spec') <- NULL 

zim_hh <- import(here("data", "raw", "PHIA data", 
                      "Zimbabwe", "zimphia2015hh.csv"))
zim_hh$centroidid <- str_squish(zim_hh$centroidid)
attr(zim_hh, 'spec') <- NULL 

zim_shp <- st_read(here("data", "raw", "shapefiles", 
                        "Zimbabwe","zwe_admbnda_adm2_zimstat_ocha_20180911", 
                        "zwe_admbnda_adm2_zimstat_ocha_20180911.shp")) 

zim_shp <- zim_shp |> 
    mutate(AREA_SQKM = units::set_units(st_area(zim_shp), km^2)) |> 
    dplyr::select(ADM0_EN, ADM1_EN, ADM2_EN, AREA_SQKM)

```

Number of polygons (units of analysis)

```{r}
nrow(zim_shp)
```

Look at centroid data

```{r}
glimpse(zim_centroids)
```

Convert point data into spatial data points

```{r}
zim_centroids <- zim_centroids |> 
    rename(x = longitude, y = latitude )
zim_centroids <- st_as_sf(x = zim_centroids, 
                          coords = c("x", "y"),
                          crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
```

Look at number of households per centroid

```{r}
zim_hh_c <- zim_centroids |> inner_join(zim_hh, by="centroidid") |> 
    group_by(centroidid) |> 
    mutate(num_hh_c = n()) |> 
    dplyr::select(centroidid, num_hh_c) |> 
    distinct()

summary(zim_hh_c$num_hh_c)
```

Number of centroids per polygon

```{r}
zim_shp$num_cent <- lengths(st_intersects(zim_shp, zim_centroids))

summary(zim_shp$num_cent)
```

Combine data from households with polygon to look at number of households per polygon

```{r}
zim_shp_hh <- st_intersection(zim_shp, zim_hh_c) # keep data from polygon and points
glimpse(zim_shp_hh)

zim_shp_hh <- zim_shp_hh |> 
    group_by(ADM2_EN) |> 
    mutate(num_centroid_polygon = n(),
           num_hh_polygon = sum(num_hh_c))

summary(zim_shp_hh$num_hh_polygon)
```

Area of polygons and number of households per km\^2

```{r}
summary(zim_shp$AREA_SQKM)

zim_shp_hh <- zim_shp_hh |> 
    mutate(num_hh_sqkm = num_hh_polygon/AREA_SQKM)

summary(zim_shp_hh$num_hh_sqkm)
```

Map of centroids per polygon

```{r}
ggplot(zim_shp) + geom_sf(aes(fill=num_cent)) + scale_fill_viridis_c()
```

## Zambia

Import the data

```{r}
#| echo: FALSE
zam_centroids  <- import(here("data", "raw", "PHIA data", 
                              "Zambia", "zamphia2016centroids.csv"))
attr(zam_centroids, 'spec') <- NULL 

zam_shp <- st_read(here("data", "raw", "shapefiles", 
                        "Zambia","zmb_adm_ecz_20230629_em_shp", 
                        "zmb_admbnda_adm3_ecz_20230629_em.shp")) |> 
    dplyr::select(ADM0_EN, ADM1_EN, ADM2_EN, ADM3_EN, AREA_SQKM)

glimpse(zam_shp)
```

Number of polygons

```{r}
nrow(zam_shp)
```

Look at centroid data

```{r}
zam_centroids <- zam_centroids |> 
    rename(x = longitude, y = latitude )
```

Convert point data into spatial data points

```{r}
zam_centroids <- st_as_sf(x = zam_centroids, 
                          coords = c("x", "y"),
                          crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
```

See how many centroids per polygon

```{r}
zam_shp$num_cent <- lengths(st_intersects(zam_shp, zam_centroids))

summary(zam_shp$num_cent)
```

Area of polygons

```{r}
summary(zam_shp$AREA_SQKM)
```

Number of centroids per polygon

```{r}
ggplot(zam_shp)  +
    geom_sf(aes(fill=num_cent)) + scale_fill_viridis_c()
```

Number of households per centroid

```{r}
zam_hh <- import(here("data", "raw", "PHIA data", "Zambia", "zamphia2016hh.csv"))
attr(zam_hh, 'spec') <- NULL 

zam_hh_c <- zam_hh |> 
    group_by(centroidid) |> 
    mutate(num_hh = n()) |>
    dplyr::select(centroidid, num_hh) |> 
    distinct()
    
```

## Malawi

Import the data

```{r}
mlw_centroids  <- import(here("data", "raw", "PHIA data", "Malawi", "mphia2015centroids.csv")) 
attr(mlw_centroids, 'spec') <- NULL 

mlw_hh <- import(here("data", "raw", "PHIA data", "Malawi", "mphia2015hh.csv"))
attr(mlw_hh, 'spec') <- NULL 

mlw_hh_c <- mlw_hh |> 
    group_by(centroidid) |> 
    mutate(num_hh = n()) |>
    dplyr::select(centroidid, num_hh) |> 
    distinct()

mlw_shp <- st_read(here("data", "raw", "shapefiles", 
                        "Malawi","mwi_adm_nso_hotosm_20230405_shp", 
                        "mwi_admbnda_adm2_nso_hotosm_20230405.shp"))

mlw_shp <- mlw_shp |> 
    mutate(AREA_SQKM = units::set_units(st_area(mlw_shp), km^2)) |>
    dplyr::select(ADM0_EN, ADM1_EN, ADM2_EN, AREA_SQKM)
                                
glimpse(mlw_shp)
```

Number of polygons

```{r}
nrow(mlw_shp)
```

Look at centroid data

```{r}
mlw_centroids <- mlw_centroids |> 
    rename(x = longitude, y = latitude )
```

Convert point data into spatial data points

```{r}
mlw_centroids <- st_as_sf(x = mlw_centroids, 
                          coords = c("x", "y"),
                          crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
```

See how many centroids per polygon

```{r}
mlw_shp$num_cent <- lengths(st_intersects(mlw_shp, mlw_centroids))

summary(mlw_shp$num_cent)
```

Area of polygons

```{r}
summary(mlw_shp$AREA_SQKM)
```

Number of centroids per polygon

```{r}
ggplot(mlw_shp) + geom_sf(aes(fill=num_cent)) + scale_fill_viridis_c()
```

Number of households per centroid

```{r}
summary(mlw_hh_c$num_hh)
```

Number of households per polygon
